A Linked Open Social Dataset for Scientific Benchmarking

### 2019-06-20 paper.tex
adicionada tabela de referencias do fb
adiantado um pouco do artigo

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black},
    pdfborder={0 0 0}
}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{changepage}
\usepackage{longtable}
\usepackage{tabularx}
% \usepackage[showframe=true]{geometry}
%
\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
\newcommand{\h} {\hphantom ~ }
% \newcommand{\textttt}[1] {\mbox{\texttt{\footnotesize#1}}}
% \newcommand{\textttt}[1] {
% \begin{verbatim} #1 \end{verbatim}
% }
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother
\begin{document}
%
\begin{frontmatter}
%
\title{A Linked Open Social Database for Scientific Benchmarking}
%
\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
\ead{fabbri@usp.br}
%
\author[pwr]{Osvaldo Novais de Oliveira Junior\fnref{kio-url}}
\ead{chu@ifsc.usp.br}
%
\cortext[corresponding]{Corresponding author}
\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
University, Brazil}
%
\fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
%
\begin{abstract}
The fields of social network analysis and complex networks
are widely researched.
Recently, a myriad of results have been reported which are based in
diverse datasets most often not accessible to researchers other than the publishing authors.
This work exposes an open database with diverse provenance and oriented
to furnish the scientific community with a friendly and common repertoire.
Current data was obtained from Facebook, Twitter, IRC, Email and the
detached instances of ParticipaBR, AA and Cidade Democr\'atica.
These were represented as linked data to homonenize access,
conform to current best practices and ease analyzes which integrate third
party and provided instances.
This document presents an outline and overall statistics of the given
database which should favor subsequent work.
\end{abstract}
%
\begin{keyword}
Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email,
Complex Networks, Text Mining
%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
\end{keyword}

@@

\section{Introduction}
In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
both research and application realms.
However, there is a lack of datasets for benchmarking results,
specially in the complex networks field, yielding diverse results from 
poorly related data.

The enormity of the digital data propels a rapid development of analysis methods
from different perspectives.
However, the used datasets in the literature differ within the scope of each research with
scarce and historical exceptions such as the
karate club dataset~\cite{newmanBook}.
On the other hand, the available linked data is not 
stable or rigorous enough to be
a public reference on statistical physics and social networks research.

This work presents a linked open social database (LOSD) with diverse
provenance, including Facebook, Twitter, IRC, Email and detached
instances.
Such data is proposed as a common repertoire for scientific
research involving networks and textual content.

@@

Integration and uniformity of access is obtained through linked data
representation, as exposed in Section~\ref{queries}.
This section introduce the underlying data in very concise terms.
One should access the Supporting Information document for
a thorough presentation of the database.

\subsection{Facebook data}
Friendship ego networks (networks whose constituents are friends of an user)
were donated from individual users in 2013 and 2014.
Friendship and interaction networks from groups were gathered from
groups where the first author was a participant.
Additionally, some groups have post texts along some metadata, such as
the number of likes.

\subsection{Twitter data}
Tweets were gathered through the Twitter streaming public API.
Each snapshot is unified by a distinct hashtag.
Edges are canonically yield by retweets but replies and user mentions
are also kept in the database.

\subsection{IRC data}
Public IRC logs were used to render IRC snapshots.
The database has records of users to which the message is directed to or
mentions.

\subsection{Email data}
Email snapshots refer to individual email lists.
All messages were obtained from the Gmane public email database~\cite{gmane}.
Each message has the original text and the text without some of the lines
from previous messages or that are software code.
Most importantly, each message instance holds the ID of the message it is
a reply to, if any.

\subsection{ParticipaBR data}
The ParticipaBR is a Brazilian federal platform for social participation.
Texts are derived from blog posts and networks are derived from
friendship and interaction criteria.

\subsection{Cidade Democr\'atica data}
Cidade Democr\'atica is a Brazilian civil society social participation portal.
Data gathered is complex in the number of types of instances.

\subsection{AA data}
The Algorithmic Autoregulation~\cite{aa} is a software development
methodology based on testifying and sharing ongoing work.
The data was gathered from different versions of the system and from an IRC
log.

\section{Methods}
Data is represented as linked open data through
RDF and ontologicaly described through a data-driven ontology synthesis
method.

\subsection{Linked open data}
Linked data refers to data published in the web in such a way that it is
machine readable and conforms to a set of best practices.
The web of data is constructed with documents on the web 
such as the web of HTML documents.
In practice, the idea of linked data can me summarized
by 1) the use of RDF to publish data on the web and 2) the use of RDF
links to interlink data from different sources.
The web is expected to be interconnected and to grow by the systematic application of four
steps~\cite{lee1}:
\begin{itemize}
    \item Use URIs to identify things~\cite{uri}.
    \item Use HTTP URIs.
    \item Provide useful information when an URI is accessed via HTTP.
    \item Provide other URIs in the description of resources so human
        and machine agents can perform discovery.
\end{itemize}

The Linked Open Data~\cite{lod} builds an ever growing cloud of data,
the global data space, which is usually
conceived as centered around the DBPedia, a linked data representation
of data from Wikipedia~\cite{dbpedia0,dbpedia}.

\subsection{RDF}
The Resource Description Framework (RDF), a W3C
recommendation, is a model for data
interchange.
It is based on the idea of making statements about resources in the form
of triples, i.e. expressions in the form ``subject - predicate -
object''.
RDF can be serialized in several file formats, including RDF/XML,
Turtle and Manchester, all which, in essence, represent a labeled and
directed multi-graph.
RDF may be stored in a type of database called a triplestore~\cite{rdf}.

As an example of an RDF statement, the following triple in the Turtle
format asserts that ``the paper has color white'':\\
\texttt{http://example.org/Things\#Paper http://example.org/hasColor\\
http://example.org/Colors\#White .}

\subsection{Data-driven ontology synthesis}
OWL Ontologies are critical tools to describe taxonomies and the
structure of knowledge.
Most ontologies are created by domain experts even though the data they
arrange is often given by a software system and has a predefined
structure.

@@

the ontological structure in data with

@@

Supporting Information document.

@@

        % \textttt{SELECT DISTINCT ?class WHERE { ?s a ?class }}
        \textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri \}}
%    \item For each class, get distinct subject classes and predicates where the
%        the object is an instance of the class:\\
%        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
%        a ?cs . \}}
% \item For each class, obtain distinct predicates (\textttt{?p}) and object classes or
% 	datatypes (\textttt{?co}, \textttt{?do}) where the subject is an instance (\textttt{?i}) of such class:\\
% 	\textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ \\
% 		\h ?i a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \\
% \}}

@@

	\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o . \}}\\

@@

	the predicate is the property (\textttt{property\_uri}).
	If the numbers match, there is an existential restriction for the class. The queries are:\\
	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
	\h ?s a <class\_uri>. ?s <property\_uri> ?o .\\ \}}
\item Find the number of instances which are subjects of triples where the predicate is the property but are not instances of the class.
	If there is zero of such instances, there is an universal restriction:\\
	\textttt{SELECT (COUNT(DISTINCT ?s)=0 as ?cs) WHERE \{\\
	\h ?s <property\_uri> ?o . ?s a ?ca . FILTER(str(?ca) != 'class\_uri')\\ \}}
\item To keep a record of the restrictions (and occurring triples), get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
	\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
		\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
\}}

@@

        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}

@@

        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s

@@

    \item For each property, find the incident range and domain with the

@@

        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
			\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\\}} \\

@@

        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}

@@

\item Render diagrams as exposed in the next section and in the Supporting Information file.

@@

\section{Results}
\label{outline}
Current overall results concern data selection and preparation for knowledge discovery.
The main result is the data made available, which enables benchmarking of scientific results
and easy experimentations.
Secondary results include data outline through figures and tables,
software support and example SparQL queries.

\subsection{Standardization}
The data is embedded into standard URIs and triples, i.e. translated to RDF.
URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
which are identified herein with the prefix \textttt{po:}.
Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
All class instances, such as participants, messages, friendships and
interactions, are linked to
snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
More specialized predicates are used for delivering text when necessary,
such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
for ParticipaBR articles.
A participant URI is unique throughout the provenance (e.g. the same for
the same participant in all Twitter snapshots).
To enable annotations which differ when the snapshot changes,
\texttt{po:Observation} class instances are used in the triple
\textttt{<participant\_uri> po:observation <observation\_uri>}.
The observation instances are then linked to the snapshot and the
data.

Instances are built on top of the class they derive from plus a hashtag character,
a provenance string (e.g. \textttt{facebook-legacy} or
\textttt{participabr-legacy}) of the snapshot they refer to, and an identifier;
i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
\textttt{irc-legacy-labmacambira} or
\textttt{email-legacy-linux.audio.devel1-20000}.

@@

\input{tables/nsnapshots}
% stats:
%% number of triples
%% number of edges
%% number of chars
%% number of users
% diagrams in the supporting information file
\subsection{Software tools}
The database is released with software for rendering itself, analyses and
multimedia artifacts.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for translating data to RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the database.}\label{tab:provenance}
\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through Gmane public database & Gmane~\cite{gmane} \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
\end{tabular}\end{center}
\end{table*}                    
\subsubsection{Topological and textual analysis}\label{ana}
Routines are available for taking topological and textual measures from
the database.
Auxiliary routines, such as performing principal component analysis
and taking Kolmogorov-Smirnov measures, are available
to ease pattern recognition.
Single, timeline and multi-scale analyzes are automated.

>>>>

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[dvipsnames]{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black},
    pdfborder={0 0 0}
}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{changepage}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{listings}
\lstdefinelanguage[]{spq}
{
    morecomment=[s][\color{ForestGreen}\bf]{<}{>},
    basicstyle = \ttfamily\footnotesize,
    commentstyle=\color{green},
    keywordstyle=\color{Plum},
    morecomment=[s][\color{Blue}\bf]{?}{\ },
    morekeywords={BASE,PREFIX,SELECT,DISTINCT,CONSTRUCT,DESCRIBE,ASK,%
    FROM,NAMED,WHERE,ORDER,BY,ASC,DESC,LIMIT,OFFSET,OPTIONAL,%
    GRAPH,UNION,FILTER,a,STR,LANG,LANGMATCHES,DATATYPE,BOUND,%
    isIRI,isURI,isBLANK,isLITERAL,REGEX,true,false},%
    sensitive=false,%
    % morecomment=[l]\#,%
    morestring=[d][\color{red}]',%
    morestring=[d]"%
}
% \usepackage[showframe=true]{geometry}
%

\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
\newcommand{\te}[1] {\texttt{\footnotesize#1}}
\newcommand{\h} {\hphantom ~ }
% \newcommand{\textttt}[1] {\mbox{\texttt{\footnotesize#1}}}
% \newcommand{\textttt}[1] {
% \begin{verbatim} #1 \end{verbatim}
% }
\linespread{1}
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}
\usepackage{booktabs}
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother
\begin{document}
%
\begin{frontmatter}
%
\title{A Linked Open Social Dataset for Scientific Benchmarking}
%
\author[pws]{Renato Fabbri\corref{corresponding}}
\ead{renato.fabbri@gmail.com}
%
\author[pwr]{Osvaldo Novais de Oliveira Junior}
\ead{chu@ifsc.usp.br}
%
\cortext[corresponding]{Corresponding author}
  \address[pws]{Institute of Mathematics and Computer Sciences, University of São Paulo (ICMC/USP), Brazil}
\address[pwr]{S\~ao Carlos Institute of Physics, University of  São Paulo
(IFSC-USP), Brazil}
%
\begin{abstract}
The fields of social network analysis and complex networks
are widely researched.
In fact, a myriad of results have been reported which are based in
diverse data, most often not accessible to researchers other than the publishing authors.
This work presents an open dataset with diverse provenance of 
human social online networking 
  and oriented
to furnish the scientific community with a friendly and common repertoire.
Current data was obtained in 2013-16 from well-known online social networking platforms:
Facebook, Twitter, IRC, Email.
Data from ParticipaBR, AA, and Cidade Democrática were also added as samples of
less known instances.
Authors recently recognized that no similar dataset was made available,
and the potential benefit of the content for the research community.
The linked data representation was adopted in order to
comply with current best practices, homogenize access, enable discovery,
and facilitate analyzes which integrate third
party and provided instances.
Futhermore, the method for obtaing and making the data available is potentially novel, and may of use for other research groups.
This document presents an outline and overall statistics of the given
dataset, which should favor subsequent work in complex and social networks,
  and in enhancing the available linked and social data.
\end{abstract}
%
\begin{keyword}
Big Data, Dataset, Benchmark Data, Facebook, Twitter, IRC, Email,
Complex Networks, Social Networks, Text Mining
%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
\end{keyword}

@@

\section{Introduction}
% In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
% both research and application realms.
The research on human social and complex networks often realy on data that express
real phenomena~\cite{c1,c2,c3}.
Such data is often not made publicly available, which hinders validation of the results and does not favor makes subsequent work.
Furthermore, there is a lack of open datasets for benchmarking results,
specially associated to the complex and social networks field,
yielding diverse results from poorly related sources.
Scarce and historical exceptions are the karate club or the dolphin datasets~\cite{newmanBook,dolphin}.
For expressing the information,
open machine-readable content and tools are desired for enabling
data linkage and discovery~\cite{o1,o2},
and the favoured protocol is thus RDF ``linked data''.
Social networks linked data is currently not stable or rigorous enough to be
a public reference on statistical physics and social networks research,
with scarce exceptions~\cite{l1,l2}
This work presents a collection of linked and open data
about online human social networks,
named LOSD (Linked Open Social Data).
This has been achieved by linking social data into a dataset using diverse
provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
Such data is proposed as a common and public repertoire for scientific
research involving networks and textual content.

\subsection{Related work}
Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
on specialized communities such as related to research and movies~\cite{s1,s2,s3,s4,s5},
or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
Contributions which are most similar to LOSD are mentioned in this section.
In~\cite{fb1}, data obtained only from Facebook is reported.
With such a focus, the authors make available a number of statistics,
but the data is offline at least since 2010.
Linked human social network data is scarcely found in the scientific community.
In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud, and made available a dataset from the Advogato platform,
dedicated to free software developers linked by trust relatioships.
There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
most often with a specific focus (users, tweets, hashtags, etc)
and not presented to the scientific community by means of a research article.
In summary, no human social network dataset presented to the
scientific community containts the diverse provenance used in LOSD,
potentially because researchers did not have the means to acquire and
share the data described in Section~\ref{acq}.

@@

\section{Methods}
LOSD is expressed as linked data through RDF
and is ontologicaly described through the
data-driven ontology synthesis method presented in Section~\ref{ont}.

\subsection{Data acquisition}\label{acq}
Research involving human subjects raise well-known ethic issues~\cite{}.
Furthermore, the online social networking platforms often restrict access
to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
In this context, research using data, e.g. from Facebook, is considered
problematic if not prohibited.
We reached a potential solution to this issue through continuated research with
collaborators in the fields of phylosophy, social sciences, and psychology.
The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
consists in researching first the social networks related to the researchers,
and keeping the procedures, data, algorithms and results publicly available.
This practices enabled collecting and sharing the data in LOSD.
Notice that, although Twitter data is available through public APIs, the other sources are more problematic.

% download via netviz when fb allowed it, donations
% twitter, bots, gmane
% contact with developers of AA, PBR, CD
% Anthropological physics

\subsection{Linked open data}
Linked data refers to data published in the web in such a way that it is
machine readable and conforms to a set of best practices.
The yielded web of data is composed of documents on the web
such as the web of HTML documents.
In practice, the publication of linked data can me summarized
by 1) the use of RDF to publish data on the web and 2) the use of RDF
links to interlink data from different sources~\cite{ld1,ld2}.
The web is expected to be interconnected and to grow by the systematic application of four
steps~\cite{lee1}:
\begin{itemize}
    \item Use URIs to identify things~\cite{uri}.
    \item Use HTTP URIs.
    \item Provide useful information when a URI is accessed via HTTP.
    \item Provide other URIs in the description of resources so human
        and machine agents can perform discovery.
\end{itemize}

The Linked Open Data~\cite{lod} is a constantly growing cloud of data,
the global data space, which is usually
conceived as centered around the DBPedia, a linked data representation
of data from Wikipedia~\cite{dbpedia0,dbpedia},
and sometimes called the Giant Global Graph (GGG, akin to the WWW).

\subsection{RDF}
The Resource Description Framework (RDF), a W3C
recommendation, is a model for data
interchange~\cite{w3cr}.
It is based on the idea of making statements about resources in the form
of triples, i.e. expressions in the form ``subject - predicate -
object''.
RDF can be serialized in several file formats, including RDF/XML,
Turtle and Manchester, which all, in essence, represent a labeled and
directed multi-graph.
RDF may be stored in a type of database called a triplestore~\cite{rdf}.

As an example of an RDF statement, the following triple in the Turtle
format asserts that ``the paper has color white'':\\
\texttt{http://example.org/Thing\#Paper http://example.org/hasColor\\
http://example.org/Color\#White .}

\subsection{RDFS and OWL ontologies}\label{sont}
One of the most important notions of the semantic web is that of an ontology~\cite{ont}.
An ontology, in this context, is a formalized conceptualization, comprised by
concepts and relations between the concepts and between the relations themselves.
In current semantic web, most simple ontologies are written using the RDFS protocol,
by which one can specify, among other things~\cite{rdfs}:
\begin{itemize}
  \item concepts;
  \item properties, which are concepts that are used as predicates and thus relate concepts;
  \item special relations between concepts that state that one concept is more general than the other\footnote{This
          kind of relation is called hypernymy. Examples: mammal is a hypernym of monkey, drink is a hypernym of beer.};
  \item the subjects and objects that can occurs in a triple where a specific property is a predicate.
\end{itemize}

One can also write an ontology using the OWL protocol,
with which all the expressive capabilities of RDFS are
available, but one can also, among other things~\cite{owl}:
\begin{itemize}
  \item state ``property axioms'', i.e. specify if a property is e.g. reflexive or transitive;
  \item state ``class restrictions'', i.e. specify if a class instance e.g. necessarily holds a relation to another class instance or data.
\end{itemize}
OWL has a richer vocabulary than RDFS and is (way more) complex.
This complexity is a drawback together with the greater computational cost
for performing inference.
The advantage is the greater power to represent conceptualizations.
Using ontologies enables automated reasoning.
For example, if a property \textttt{:r} is known to relate only mokeys to trees,
then, if there is a triple \textttt{:a :r :b}, the instance \textttt{:a} is considered a monkey,
and instance \textttt{:b} is considered a tree,
even if there is no explicit declaration similar
to \textttt{:a rdf:type :Mokey}.

For the data described in this article, the following features are most relevant~\cite{rdfs,owl}:
\begin{itemize}
  \item Using RDFS, one may define a property domain, i.e. the class instances that are allowed as the object in a triple with the property.
  \item Using RDFS, one may define a property range, i.e. the class instances that are allowed as the predicate in a triple with the property.
  \item Using OWL, one may define an existential restriction for a class \textttt{:C},
    i.e. declare that an instance \textttt{:C\#I} of such class has at least one relation with a property \textttt{:p} and a class \textttt{:C2} in the form
    \textttt{:C\#I :p :C2\#I}, where \textttt{:C2\#I} is an instance of the class \textttt{:C2}.
  \item Using OWL, one may define an universal restriction for a class \textttt{:C},
    i.e. declare that any relation of an instance \textttt{:C\#I} of such class, with a property \textttt{:p} of the form \textttt{:C\#I :p :CX\#I},
    implies that \textttt{:CX\#I} is an instance of a specific class, e.g. \textttt{:C2}.
\end{itemize}

\subsection{Data-driven ontology synthesis}\label{ont}
OWL Ontologies are tools to describe taxonomies and the
structure of knowledge~\cite{owl1}.
Most ontologies are created by domain experts even though the data they
arrange is often given by a software system and has a predefined
structure~\cite{owl2}.

@@

the ontological structure in the data with

@@

Supporting Information document of this article.

@@

        \h\textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri .\}}

@@

	\h\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o .\}}\\

@@

        the predicate is the property (\textttt{property\_uri}) and the object is
        an instance of the same class (\textttt{class2\_uri}).
	The queries are:\\
	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
        \h\textttt{SELECT DISTINCT ?oc WHERE \{\\
	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a ?oc . \\ 
        \h\}}\\
        Then, for each object class (\textttt{object\_curi}) in \textttt{?oc}:\\
	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs2) WHERE \{\\
	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a <object\_curi> .\\         \h\}}\\
        When \textttt{?cs1} matches \textttt{?cs2}, an existential restriction is found. And with the query:
	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
	\h\h ?s a <class\_uri> . ?s <property\_uri> ?o . ?o a ?oc .\\
        \h\h FILTER(str(?oc) != 'object\_curi')\\
        \h\}}\\
       When \textttt{?cs} is 0, a universal restriction is found. 
\item To keep a record of further restrictions, get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
	\h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
        \h\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
        \h\}}

@@

        \h\textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}

@@

        \h\textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s

@@

          \item For each property, find the incident range (possible subjects) and domain (possible objects) with the

@@

        \h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
        \h\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
        \h\}} \\

@@

        \h\textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}

@@

      \item Render diagrams to achieve figures as exemplified in Section~\ref{sdia} and in the Supporting Information document.

@@

Integration and uniformity of access is obtained through linked data
representation, as described in Section~\ref{queries}.
This section introduces the underlying data in very concise terms.
One should access the Supporting Information document for
a more information about the dataset.

\subsection{Facebook data}
Friendship ego networks (networks whose nodes represent friends of a user)
were donated from individual users in 2013 and 2014.
Friendship and interaction networks from groups were gathered from
groups where the first author was a participant.
Additionally, some groups have post texts along some metadata, such as
the number of likes.

\subsection{Twitter data}
Tweets were gathered through the Twitter streaming public API.
Each snapshot is unified by a distinct hashtag.
Network links are canonically yield by retweets,
but replies and user mentions are also kept in the database.

\subsection{IRC data}
Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
The database has records of relations yield by directed messages and
mentions.

\subsection{Email data}
Email snapshots refer to individual email lists.
All messages were obtained from the Gmane public email archive~\cite{gmane}.
Each message has the original text and the text without some of the lines
from previous messages or that are software code or author signature.
Most importantly, each message instance holds the ID of the message it is
a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}

\subsection{ParticipaBR data}
ParticipaBR was a Brazilian federal platform for social participation,
the most important platform from the myriad of social participation platforms
that were available in 2011-2018~\cite{spbr}.
Texts are derived from blog posts and networks may be obtained from
friendship and interaction criteria.

\subsection{Cidade Democr\'atica data}
Cidade Democrática is a Brazilian civil society social participation portal.
Data gathered is complex in the number of types of instances.

\subsection{AA data}
AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
it is based on testifying and sharing ongoing work.
The data was gathered from different versions of the system and from an IRC
log.

\section{Results}
\label{outline}
Current overall results concern data selection and preparation for data discovery.
The main result is the data made available, which enables benchmarking of scientific results
and easy experimentations, and no compliant dataset was found by the authors.
Secondary results include data outline through figures and tables,
software support and example SparQL queries.

\subsection{Standardization}
The data is embedded into URIs and triples, i.e. translated to RDF.
URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
which are identified herein with the prefix ``\textttt{po:}''.
Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
All class instances, such as participants, messages, friendships and
interactions, are linked to
snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
More specialized predicates are used for delivering text when necessary,
such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
for ParticipaBR articles.
A participant URI is unique throughout the provenance (e.g. the same for
the same participant in all Twitter snapshots).
To enable annotations which differ when the snapshot changes,
\texttt{po:Observation} class instances are used in the triple
\textttt{<participant\_uri> po:observation <observation\_uri>}.
The observation instances are then linked to the snapshot and the
data.

Instance URIs are built using the URI from the class they derive from plus a hashtag character,
a provenance string (e.g. \textttt{facebook-legacy} or
\textttt{participabr-legacy}), and an identifier;
i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
\textttt{irc-legacy-labmacambira} or
\textttt{email-legacy-linux.audio.devel1-20000}.

@@

\input{tables/nsnapshots}
% stats:
%% number of triples
%% number of edges
%% number of chars
%% number of users
% diagrams in the supporting information file
\subsection{Software tools}
The authors made available software for rendering analyses and
the data set itself.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for expressinf the data as RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
  {\renewcommand{\arraystretch}{1.2}
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the database.}\label{tab:provenance}
\begin{tabular}{ l | p{3cm} p{3cm} c }
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through the Gmane public archive & Gmane~\cite{gmane} \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\
\end{tabular}\end{center}
\end{table*}                    
  }
\subsubsection{Topological and textual analysis}\label{ana}
Routines are available for taking topological and textual measures from
the LOSD data.
  Auxiliary routines, such as for performing principal component analysis (PCA)
  and obtaining statistics derived from the Kolmogorov-Smirnov test~\cite{kol}, are available to ease pattern recognition.

% \subsubsection{Multimedia rendering}\label{media}
% It is a core purpose of framework to provide routines for rendering
% audiovisualizations of the data.
% Social structures are rendered into music, images and video animations
% through the Percolation toolbox~\cite{percolation} in association with
% the Music and Visuals toolboxes~\cite{music,visuals}.
